---
title: "Project 2"
author: "Matthew Farris"
date: "October 11, 2015"
output: 
  html_document:
    toc: true
---
# Does the Power of the Pope Make Me a Better Player? 

For this assignment, I decided to go a little off the beaten path. As we are still in the introduction of data acquistion, and these Projects are more to help build a basis for our future understanding of data management,  I feel we can still create solid data and code that highlights a way of extracting data, while still having a little fun. I wanted to focus on rvest in the project, so I took the two data sets from Sanjive Kumar on Brazil's Footbal Team, and the Motocross data from Maxwell Wagner. I also noticed that the data for the Pope's Popularity by Andrew Goldberg was an interesting choice, not exactly like the other data, but an opportunity to compare data. I thought it would be fun to play around with this and maybe to answer the question:  Is there any correlation, between the Pope's Popularity and a Team's winning streak? 


## Data Extraction

For the first part of this project, we extrackted the necessary data an created 
```{r}
#Initializing Packages
library(rvest)
library(dplyr)
library(tidyr)
library(RCurl)
library(stringr)
library(ggplot2)
```


### Brazil's FIFA Record

The reason I primaily picked the Motocross and FIFA data was because it was from Wikipedia. This data can easily be extracted using the Rvest Package. I did some research into the package, and will rely heavily on one source in particular: 

http://www.r-bloggers.com/using-rvest-to-scrape-an-html-table/   

This one set of code requires one to inspect the elements of a webpage and locate the specific path in the source code. I like this way particularly, as it allows for to extract the dataset directly from the site. No need to create a csv file in another program. So, the extracted data looks as follows: 


```{r}
WC_URL <- "https://en.wikipedia.org/wiki/Brazil_national_football_team#FIFA_World_Cup"

WC_Data <- WC_URL %>% 
  html() %>% 
  html_node(xpath ='//*[@id="mw-content-text"]/table[2]') %>% 
  html_table(header = NA, trim = TRUE, fill= TRUE)
head(WC_Data)

```

So the above code works at extacting a very rough dataset, but it did extract it! For our purposes here, I was only interested in the first table (as you can see from the wikipedia Page it attached two tables into one) So first things first, clean up and tidying:

```{r}
# Remove only the first 9 Columns which we will be working with.
WC_Brazil <-WC_Data[1:9]
#Resets Header to the Second Row
names(WC_Brazil) <- c("Year","Round", "Position", "Played", "Won", "Draw", "Lost","Goals","Goals-Against")
#Removes Duplicate Header Row
WC_Brazil <- WC_Brazil[-1,]
#Removes Total at the bottom (Not Necessary)
WC_Brazil <- WC_Brazil[-21,]
```
Now we have a cleaner data frame that we can work with, but there is still some tidying that needs to be done. First, the data that we are going to primarily work with in this data set is the final position. However, this is a none numeric value, so tidyr can be used to extact this data: 

```{r}

WC_Brazil$Position <- extract_numeric(WC_Brazil$Position)

head(WC_Brazil)
```

### Motocross

For the Motocross data a similar Extraction was performed, and we created to following dataset.
```{r}
MC_URL <- "https://en.wikipedia.org/wiki/Motocross_des_Nations"

MC_Data <- MC_URL %>% 
  html() %>% 
  html_node(xpath ='//*[@id="mw-content-text"]/table[3]') %>% 
  html_table(header = NA, trim = TRUE, fill= FALSE)

#Trimmed an excess column on the end: 
MC_Data <- MC_Data[,-18]

head(MC_Data)

```
For the most part this dataset is much more tidy than the previous one, and didn't require much deletion or extraction of data. However, it did require us to gather the data into a much more manageble set. So that was the first part that we did:

```{r}
MC_Total <-MC_Data %>% gather("Year","Place",2:17,na.rm = TRUE)
#Another step needed would be get a numeric value for the postion
MC_Total$Place <- extract_numeric(MC_Total$Place)
#Last we need to omit the NA values, as they are unnecessary for our analysis downstream
MC_Total<-na.omit(MC_Total)

head(MC_Total)
```


### Pope Popularity 

For the Pope's Popularity, unfortunately, the data was in an image file. At the moment, I was unable to extract the info into a reasonable group. So I created a CSV file in the format of the data presented on the Discussion board. I kept the CSV to the format as best I could, but for simplicity I did change things a bit (I realize that this Project is to aid our knowledge of tidying data, but some things especially in a .csv file are easier to manipulate in the file itself). I tried to use each of these datasets to emphasize a particular extraction and data manipulation, for this one I used some regular expressions to extract a particular item in the file. 

```{r}
#Extracting the URL text 
popeURL <-getURL("https://raw.githubusercontent.com/mfarris9505/PopePop/master/PopePopularity.csv")

#Reading the txt file into "CSV"
pope_data <-read.csv(text = popeURL)
#Re-naming the columns
names(pope_data) <- c("Pope", "Date","Favorable", "Unfavorable", "No.Opinion")

head(pope_data) 
```

This data is pretty tidy, however, for my analysis, I want to extract the specific year when the poll was taken. So to extract the data, we used the the str_extract function. 

```{r}
pope_data$Date <-str_extract(pope_data$Date,"[:digit:]{4}")

#There were two 2005 years in this dataset for Pope Benedict and one for Pope John Paul, for ease (and because I needed a 2006 and a 2004 data point) I changed the December poll to 2006 and the Pope John Paul to 2004. We have yet to learn about the problems of changing data, so I figured this would be OK.
pope_data$Date[6] <- 2006
pope_data$Date[8] <- 2004
head(pope_data)
```

Now that we have all the data in a group we can do some analysis.

## Results and Analysis 
For the analysis, I did the bare minimum that was requested for the two threads (Motocross and Pope data), which is shown below. For the World Cup data and Motocross data, I created two datasets and combined it with Favorability of the Pope during the given years. For the Motocross data, I picked 4 countries specifically USA, Japan, Ireland, and Brazil 


### Thread Analysis

As stated, I couldn't quite extract the data for the opponents, so I did some generic plotting. First, I was interested in see if the number of goals scored vs. the position that Brazil made it too. This was accomplished as follows:  

The motocross data was interesting to look at, and using dplyr we came to some interesting finds. First the average placement of each country: 

```{r}
MC_Total %>% group_by(Team)  %>% 
  summarise(Mean = mean(Place))



```

For the Pope dataset, I created two new Columns using the Mutate function, one to obtain what he called Awareness, and the other as Net Favoablility:

```{r}
pope_data <- pope_data %>% mutate(Awareness = Favorable + Unfavorable) %>%  mutate(Net_Favor = Favorable - Unfavorable)

pope_data

```


### Combined data analysis

This is where I take leaps and bounds past normal, and took it one step further: I combined data... Why? Because I can... 

First step was the create a simple data table from the Pope data this was done as followed: 

```{r}
combo_data <- subset(pope_data,select = c(Date, Favorable))

names(combo_data) <- c("Year","Favorable")

# I mutated this into A percentage
combo_data <- combo_data %>% mutate(Favorable = Favorable/100) 

# I next pulled data fromthe World Cup:
WC_Combo <- subset(WC_Brazil, select= c(Year,Position))

# And finally pulled the Motocross Data:
MC_Japan <- MC_Total %>% filter(Team == "Japan")
MC_USA <- MC_Total %>% filter(Team == "USA")
MC_Ireland <- MC_Total %>% filter(Team == "Ireland")
MC_Brazil <- MC_Total %>% filter(Team == "Brazil")

#Removes excess Column
MC_Japan <- MC_Japan[,-1]
MC_USA <- MC_USA[,-1]
MC_Ireland <- MC_Ireland[,-1]
MC_Brazil <- MC_Brazil[,-1]

#Using the Left_join we combined each of these separate datasets into one: 
combo_data <- left_join(combo_data,WC_Combo, by = "Year")
combo_data <- left_join(combo_data,MC_Japan, by = "Year")
combo_data <- left_join(combo_data,MC_USA, by = "Year")
combo_data <- left_join(combo_data,MC_Ireland, by = "Year")
combo_data <- left_join(combo_data,MC_Brazil, by = "Year")

names(combo_data) <- c("Year","Favorable","Brazil WC","Japan MC","USA MC","Ireland MC", "Brazil MC")

combo_data
```

After all that... We successfully found absolutely nothing... I was going to plot this in a graph to determine if there was some correlation (ie. Favorability vs. Place), but frankly the dataset is a bit to sparse to do that. Furthermore, looking at the data, one can easily see that their is no relationship. However, for the sake of argument, we did plot the data for Ireland, as it looked promising: 


